---
title: "Sports Betting on Unemployment"
author: "Kevin Phan"
date: "2025-05-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(tidyverse)
library(fixest)
library(did)
library(stargazer)
library(modelsummary)
# reading in the dataset
df <- read.csv("clean_data/df_clean.csv")

```


# Basic Summary 1: Ungrouped
```{r}
summary <- df %>% select(population, unemployment, bachelors, rgdp_2017)
labels = c("Population (Thou)",
           "Unemployment (%)",
           "Bachelor's Degree (%)",
           "Real GDP (2017, Thou)")
stargazer(summary, type = "latex", title = "Descriptive Statistics of All States", digits = 1, covariate.labels = labels)
```

# Basic Summary 2: Grouped by Legal, Retail Only, Online Only
```{r}

summary_retail <- df %>% filter(retail == 1) %>% select(population, unemployment, bachelors, rgdp_2017) 

stargazer(summary_retail, type = "latex", title = "Descriptive Statistics (States Legalized Retail)", digits = 1, covariate.labels = labels)

summary_online <- df %>% filter(online == 1) %>% select(population, unemployment, bachelors, rgdp_2017)

stargazer(summary_online, type = "latex", title = "Descriptive Statistics (States Legalized Online)", digits = 1, covariate.labels = labels, notes = "Many states that have legalized online gambling have also legalized retail gambling (and vice versa). This explains why the number of observations between the second table and third table are seemingly too high.")

```

#Plot of US Map with Legalization

```{r}
library(usmap)
# must plot the states that are fully, those that are partially, and those that are not legalized

df_plot <- df %>%
  mutate(legal_status = case_when(
    (retail == 1 & online == 1) ~ 'Fully Legal',
    (retail == 1 & online == 0) ~ 'Retail Only',
    (online == 1 & retail == 0) ~ 'Online Only',
    TRUE ~ 'Illegal'
  ))
plot_usmap(color = "black", 
           regions = "states",
           data = df_plot,
           values = "legal_status") + labs(
             fill = "Legal Status"
           ) + theme(legend.position = 'bottom', plot.title = element_text(hjust = .5, size = 18, face = "bold")) + scale_fill_manual(values = c("Fully Legal" = "limegreen", "Retail Only" = "lightblue", "Online Only" = "turquoise", "Illegal" = "darkgrey"), na.translate = FALSE) +
  ggtitle("Status of Sports Betting by State")
```




# Analysis
```{r}
## modifying the unemployment rate and the bachelor's degrees rate
# adding in the time-to-treatment variable.
df_analysis <- df %>%
  filter(!(state == "Nevada"),
         date > 2015) %>% # nevada's retail vs sports implementation too extreme based on calculation.
  mutate(unemployment = unemployment/100,
         bachelors = bachelors/100,
         time_treat_retail = ifelse(retail == 1, date - retailyear, 0),
         time_treat_online = ifelse(online == 1, date - onlineyear, 0),
         time_treat = ifelse(legal == 1, date - legal.year, 0))

```

# Trends in Legalization
Plot that shows the freq of legalization.
```{r}
df_analysis %>% 
  group_by(date) %>%
  select(date, time_treat, legal) %>%
  filter(time_treat == 0) %>%
  summarize(freq = n() - 12) %>%# subtract 12 for the twelve states that have not legalized yet 
  ggplot(data = ., aes(date, freq)) +
  geom_line() +
  geom_point() + theme_classic()

```

# Pre-Trends Testing

This is somewhat useless, but interesting to look at.
```{r}

df_analysis %>% filter(time_treat_retail < 0) %>%
  feols(unemployment ~ i(time_treat_retail, retailyear) + population + rgdp_2017 + bachelors | abbrev + date,
        cluster = ~ abbrev,
        data = .) -> model_pt_r

iplot(model_pt_r, main = "Pre-trend Unemployment (TWFE)", xlab = "Year to Legalization")

summary(model_pt_r)

df_analysis %>% filter(time_treat_online < 0) %>%
  feols(unemployment ~ i(time_treat_online, onlineyear) + population + rgdp_2017 + bachelors | abbrev + date,
        cluster = ~ abbrev,
        data = .) -> model_pt_o

iplot(model_pt_o, main = "Pre-trend Unemployment (TWFE)", xlab = "Year to Legalization")

summary(model_pt_o)

```

# TWFE model & SA model
The models as well as their tables/figures.
```{r}
model_twfe1 <- feols(unemployment ~ i(time_treat_retail, retail, ref = -1) + 
                      population + rgdp_2017 + bachelors | abbrev + date,
                    cluster = ~ abbrev,
                    data = df_analysis)


summary(model_twfe1)

model_sa1 <- feols(unemployment ~ sunab(retailyear, date) + 
                      population + rgdp_2017 + bachelors | abbrev + date,
                    cluster = ~ abbrev,
                    data = df_analysis)

summary(model_sa1)
iplot(model_twfe1, main = "Retail Gambling on Unemployment (TWFE)") 
iplot(model_sa1, main = "Retail Gambling on Unemployment (Sun & Abraham)", xlab = "Year to Implementation")

model_twfe2 <- feols(unemployment ~ i(time_treat_online, online, ref = -1) + 
                      population + rgdp_2017 + bachelors | abbrev + date,
                    cluster = ~ abbrev,
                    data = df_analysis)

model_sa2 <- feols(unemployment ~ sunab(onlineyear, date) + 
                      population + rgdp_2017 + bachelors | abbrev + date,
                    cluster = ~ abbrev,
                    data = df_analysis)


summary(model_twfe2)
summary(model_sa2)

etable(list(model_sa1, model_sa2), tex = T)
iplot(model_twfe2, main = "Online Legalization Effect on Unemployment (TWFE)", xlab = "Year to Legalization", sub = "Note: Reference line is one year prior to legalization.")
iplot(model_sa2, main = "Online Legalization Effect on Unemployment (Sun & Abraham)", xlab = "Year to Legalization")


agg_cohort_sa1 <- aggregate(model_sa1, agg = c("cohort"))
agg_ATT_sa1 <- aggregate(model_sa1, agg = "ATT", use_weights = T)
agg_cohort_sa2 <- aggregate(model_sa2, agg = c("cohort"))
agg_ATT_sa2 <- aggregate(model_sa2, agg = "ATT", use_weights = T)

```

## Supplemental Analysis
We look at credit card debt per capita. 

```{r}
library(readxl)
credit_debt <- read_excel("~/Desktop/Math Programming/ECON427/final_project/other_data/debt_by_state.xlsx", 
    sheet = "creditcard", skip = 8)

credit_debt <- credit_debt %>%
  pivot_longer(
    cols = Q4_2003:Q4_2024,
    names_to = c("year"),
    values_to = "debt") %>%
  mutate(year = as.numeric(substring(year, 4, 10))) %>%
  rename("abbrev" = "state",
         "date" = "year")

df_debt <- merge(df_analysis, credit_debt, by = c("date", "abbrev"))

```

```{r}

# hmm, if you have fixed effects for state and time, then it fails...

df_debt %>%
  feols((debt) ~ sunab(legal.year, date) + population + rgdp_2017 + bachelors | abbrev + date, clusters = ~abbrev, data = .) -> model_debt

iplot(model_debt)
df_debt %>%
  feols((debt) ~ sunab(retailyear, date) + population + rgdp_2017 + bachelors | abbrev + date, clusters = ~abbrev, data = .) -> model_debt_r

summary(model_debt_r)
iplot(model_debt_r, main = "Retail Gambling Effect on Credit Card Debt Per Capita", xlab = "Year to Implementation")

df_debt %>%
  feols((debt) ~ sunab(onlineyear, date) + population + rgdp_2017 + bachelors | abbrev + date, clusters = ~abbrev, data = .) -> model_debt_o

summary(model_debt_o)
etable(list(model_debt_r, model_debt_o), tex = F)

iplot(model_debt_o, main = "Online Gambling Effect on Credit Card Debt Per Capita", xlab = "Year to Implementation")

agg_cohort_debtr <- aggregate(model_debt_r, agg = c("cohort"))
agg_ATT_debtr <- aggregate(model_debt_r, agg = "ATT", use_weights = T)

agg_cohort_debto <- aggregate(model_debt_o, agg = c("cohort"))
agg_ATT_debto <- aggregate(model_debt_o, agg = "ATT", use_weights =T)

```




```{r}
# cohort average treatment effects and the general ATT
att_labels = c("Retail on Unemployment", "Online on Unemployment", "Retail on Credit Debt", "Online on Credit Debt")

ATT <- rbind(agg_ATT_sa1, agg_ATT_sa2, agg_ATT_debtr, agg_ATT_debto)
stargazer(ATT, title = "ATT from Diff-in-Diff Models", type = "latex", notes = "The ATT is calculated from the CATT using weighted averages. These weighted averages are calculated under the hood when we run Sun and Abraham's estimator using the fixest package in R.")


cohorts <- as.data.frame(rbind(agg_cohort_sa1, agg_cohort_sa2, agg_cohort_debtr, agg_cohort_debto))
# stargazer(cohorts, main = "Cohort-Level Effects (By Year) By Model") no work

kableExtra::kable(cohorts, type = "text", digits = 3, main = "Cohort-Level Effects (By Year) By Model") %>% kableExtra::kable_classic()
```





